\section{Gamification}

Kapp defines gamification as “Using game-based mechanics, aesthetics and game thinking to engage people, motivate action, promote learning, and solve problems”. There are many definitions of gamification but the one that is always similar is that it is a tool to “increase engagement in some activity using game features, providing enjoyment and fun”, the use of game elements in non-gaming systems to improve user experience and user engagement \cite{karlmkapp2012}. Unsurprisingly, there has been a growing trend of the use of gamification to capture the target audience of traditional applications \cite{luisRodrigues2016}. One of the benefits that gamification brings is the promise of new design patterns and a more interactive user experience. With the interests of businesses to keep users engaged and entertained in using the app for as long as possible, it is not surprising that we have seen the application of game mechanics of many non game related contexts.

In gamification, the focus is not the game itself but they have a significant role as a motivation tool for users to continue using the app \cite{alidarejeh2016}. Gamification has seen a wide array of applications in different disciplines. Ali Darejeh and Siti Salwah Salim lists down the most common contexts of the usage of gamification:
\begin{itemize}
\item Human Resources. Many applications have proven to be successful to increase personnel motivation, engagement and working performace.
\item Health care and Sport. With the help of gamification, people have shown an increase of desire to do physical activities, which promotes to live healthy and motivate hospital patients to continue taking their medicines.
\item E-Learning. Gaming in education helps students to spend more time in learning the reading material and increase their class activity by asking and answering questions in the systems.
\item Data Collection. Gamification motivates people to participate in the collection of data, such as getting crowdsourced information for research.
\item Online Community. Gamification encourages users to to participate in a certain online community such as blogs, social websites and others.
\item Software Popularity. Lastly, gamification increase the loyalty of customers to use specific products for a long time. This also translates to an increased frequency of using the software and a discovery-based learning.
\end{itemize}

There are many examples of gamification that can be seen in present mobile applications such as Endomondo, Nike Plus and FourSquare \cite{alidarejeh2016}. In Endomondo and Nike Plus, they make use of special badges and points when they run a certain distance or participate in a predefined activity listed in the app. Similarly in Foursquare, users receive badges when they visit a special place and use the app to check-in. The use of virtual badges and points are to make the users more motivated, and to continuously encourage the use of the app. Most apps also make use of leaderboards to add a social dimension and a sense of competition among its users.

\subsection{Self Determination Theory}

The theory behind the motivation of users when using a gamified application is usually the SDT (Self Determination Theory) \cite{karlmkapp2012}. The theory states that the human being has three basic needs: competence, relatedness, and autonomy. A person is motivated to perform a certain activity based on the level of these three needs \cite{andrade2016bright}. The theory further suggests that the degree of motivation can vary from amotivation, without any motivation at all to perform an activity, to intrinsic motivation, when the user does not need any external incentives to perform the activity \cite{karlmkapp2012}. The gamification theory then proposes that the use of gaming elements in any application will be able to satisfy some of the users’ basic needs which would make the activity more attractive, even though the user is not intrinsically motivated.

As seen in many gamified apps, gamification is commonly associated with adding a scoring system in order to motivate users--badges, proof of status, points, and so on. While this may work out successfully in the short term, this may have a detrimental effect over the long-term. Scoring systems function as rewards that give users external motivation. Psychological studies have found that people respond immediately to rewards and external motivation can easily be provided. However, external motivation has an inverse relationship to internal motivation, wherein the user continues engaging in an activity because of an innate desire to; he would do it even without a reward \cite{reeve2004}.

An example of rewards decreasing internal motivation can be found in the school setting. Students’ work are graded by teachers, and these grades supposedly function as feedback on how they are doing. While this can motivate students to improve their performance, students also tend to cease to put effort when they realize that they will not be graded. This shows the pitfall of targeting users through rewards: they have to keep receiving rewards, or else engagement will drop altogether. Reffered to as reward gamification, and among the criticisms thrown at it is that it is hardly “gamelike” at all, as the most prominent game element it borrows is the scoring system and this is the least interesting part \cite{nicholson2012user}.

\subsection{Meaningful Gamification}

In order to propose an alternative model that is less exploitative of users and that has a greater possibility of sparking long-term change, an alternative view is the organismic integration theory. Organismic Integration Theory, which studies how external rewards can be internalized and integrated into one’s sense of self, states that users fare better when the activity somehow becomes aligned with their personal goals and connected with their values. In comparison, users motivated by external controls will have a more negative perception of the activity \cite{reeve2004}.
Another theory is Universal design for Learning. Courses should adapt to different learners \cite{rose2000universal}. When transferred to a gamification context, this means that either users should be allowed to make their own goals or there should be different ways for them to reach these goals. A scoreboard may appeal to the competitiveness in some users, but other users might have a different way of assessing how much they’ve achieved and thus would not find this meaningful \cite{nicholson2012user}.

Ultimately, Nicholson is putting forward a model that he calls meaningful gamification, where the user is driven by intrinsic rather than extrinsic motivation. The user must thus be at the center. Rather than asking about what benefits the organization would like to have, the designer of the application ought to be concerned first about how to help the user. In this way, the benefits become mutual: users become engaged in and derive value from the application, which of course would have a positive turnout for the organization \cite{nicholson2012user}.

However, this is not meant to discredit using reward systems altogether, as they still have their use when it comes to getting immediate responses from the user. In very specific situations, they can also pave the way for long-term change. When getting the user to learn a real-world skill, such as programming or cooking, doling out rewards could work. As the user progresses, he might eventually realize the significance of learning such a skill and derive intrinsic satisfaction without the need for external rewards anymore. This is why the recommended approach is to bring users in using a light reward system initially, then once their attention is sufficiently captured, this can be replaced by meaningful gamification \cite{scottnicholson2014}.

As for what constitutes meaningful gamification, Nicholson lists down six elements taken from game design: play, exposition, choice, information, engagement, and reflection.
\begin{itemize}
\item Play is almost always synonymous with fun; the user must participate in it by her own choice, and she should be given ample freedom to explore and make mistakes. The difference between play and a game is that “a game is a form of play with goals and structures \cite{maroneykevin2001}.” A gamified system must thus allow the users to create their own rules and change them when necessary--temporary rules, in other words.
\item Exposition refers to embedding a narrative in the system. The user can see herself as being part of a storyline, journey, or quest, which provides continuity through time and gives a sense of progress and growth. This is an especially powerful tool because human beings naturally process everything in terms of narrative.
\item If a set narrative is to be provided rather than giving the user free reign to create his own, which can be difficult to implement in technological systems, then this has to be balanced with the user’s autonomy or freedom of choice. The more that a user can customize an application, or the more tailor-fitted it is to him rather than being generic, the more involved the user feels, and the greater his sense of positive wellbeing.
\item Next is information. Rather than stopping short at being showered with points, users must also know how and why they are being rewarded. An example of this element in an actual game is the wise sage that guides the player by giving her knowledge. This caters to the user’s need for mastery, which is one of the components of intrinsic motivation.
\item Engagement can refer to either: (1) capturing the user’s interest and getting him into a state of flow; or (2) getting her to connect with others. Flow occurs when one becomes so absorbed in doing an activity that one loses track of time. This happens when an activity pushes the user out of her comfort zone, but at a level that she can handle, which leads the user to improvement. As she grows, the level of difficulty that she needs in order to experience flow also rises. If the level is kept too easy and no opportunity for growth is provided, the user lapses off into boredom \cite{mihalycsikszentmihalyi2008}. On the other hand, the social aspect of engagement is based on the user’s feeling of connection with other people, commonly expressed in applications through social networks and peer-to-peer exchanges. Avenues of social engagement can be chat boards and forums; they can also have a competitive slant, such as scoring systems or users playing against each other.
\item Finally, reflection means that the user can find a significant connection between the system and his life experiences. Without this, it would be impossible to find meaning in the system or to engage in it meaningfully. The process of reflection starts with a direct experience; the user then relates this experience with his life, then makes generalized conclusions based on this, and applies this to a whole set of new experiences as the cycle starts all over again \cite{kolb1981learning}. Through reflection, the user can take what he learned from the system and actually carry it over to the real world.
\end{itemize}

Aside from letting rewards-based transition to meaningful gamification, the gamification must also be designed to end; it should not be kept up indefinitely, as this will make the user dependent. Ideally, gamification should leave the user more equipped to deal with the real world: the application should instill long-term change in the user. As Nicholson points out, \enquote{the goal of the journey is to remove the gamification layers entirely... Gamification should not be thought of as a cycle, but as a journey to bring about lifelong change} \cite{scottnicholson2014}.
