\chapter{Introduction}

In the era of online social networking, it can be said that people are more expressive than ever since there is now an avenue to express publicly emotions (sentiments and reactions) and actions (behavior, location) in realtime \cite{deller2011twittering}. Because of this, Twitter has become a prime candidate to learn more about the daily conditions of people. The immediacy of the contents from Twitter makes it an ideal source of data for analysis in areas such as disaster controls, marketing strategization, population profiling, health emergency tracking, and more \cite{zeng2010social, ribarsky2014social}. In the context of the Philippines, a country considered to be the social networking capital of the world, the potential of Twitter posts can be actualized  as a layer to improve on existing epidemiological systems employed in the country.


% In a world where social media accounts are rapidly becoming people's virtual identity online, gathering information about a whole population can literally be as easy as a click on a mouse. Twitter is one of the top social networking websites that users visit to keep their ``friends'' informed with things happening in their life . . This research capitulates on Twitter's features as a social networking website to track posts directly or indirectly concerning the users' state of health. Personal accounts - not that of an organization such as news accounts - will be the prime focus, since this will provide the research a more population-based surveillance needed for spatio-temporal modeling. 
\textit{"Infodemiology"} is a term coined by Eysenbach when dealing with surveillance of public health-related text in the Internet \cite{Eysenbach:2009aa}. It aims to measure and track the behaviour of information online with regards to public health. Social networking websites, such as Twitter, had proved themselves to be key to creating infodemiological studies after previous successful studies \cite{nascimento2014real, Chew:2010aa, Bragazzi:2013aa,eysenbach2006infodemiology}. The primary focus of this research is the infodemiology of three (3) of the top communicable diseases - Measles, Influenza, Typhoid Fever -  in the Philippines, using Twitter \cite{dohrecord2015}. 

%In recent years, advancements have been made to track certain diseases using different health analytics tools \cite{Simpao:2014aa, Wills:2014aa}. The methodologies present in extracting data from Hospital Information Systems (HIS)

%This is not the priority for this research as these methodologies directly gathers data from hospital information systems (HIS) that are tied to hospitals' daily activities. The methodology used in the extraction of data in an HIS  will only then serve as a guide to retrieve data from social networking websites to gather relevant information for the research, and the data gathered may then be cross-referenced to actual statistical records to produce certain correlations. 

This research is aligned with the \textit{Philippine eHealth Strategic Framework and Plan 2020} that may help in strengthening existing eHealth solutions through adding to the \textit{Information Sources} pillar of the framework. Through the use of ICT, the state of eHealth in the Philippines should be improved to provide even faster universal healthcare to people. The research may pave the way for complementing the usual way of tracking diseases that usually requires work that take up a lot of time \cite{gomide2011dengue}. 

% Some methods include the use of turnaround surveys that are given to patients, and for some other diseases, such as tuberculosis, health stations that are part of a bigger health information system are set-up to aggregate data, which is the case in tracking tuberculosis in the Philippines \cite{vianzon2013tuberculosis}. 

The power of social media lies on the immediacy it has, and the amount of context a researcher could get from the strings of texts that users post. Twitter, for example, has a maximum limit of 140 characters per tweets. This limit helps in making it easier to filter subjects, sentiments, and data through the use of computational algorithms \cite{schmidt2012trending}. The extracted data may help researchers in finding out patterns in them, which may lead into creating different kinds of models based on them.

%Currently, there is no comprehensive profile done by the Philippines' Department of Health (DOH) for the Influenza cases in the country. On the other hand, the DOH publishes weekly - sometimes tri-weekly - reports on recorded Influenza-like cases through their website \cite{dohflusurvey}. In this reports, basic statistics are shown, which includes gender ratio, age groupings, case and mortality counts, and municipality segregation. These data are vital for this research as these are the ones that are cross-referenced with users' posts. 

The Department of Health (DOH) is the government entity in charge of collecting and storing statistics regarding diseases in the country. The Epidemiology Bureau (EB) is an office under the DOH whose mandate is to develop and evaluate disease surveillance systems. Currently, the data they gather are in public domain, and can be accessed easily through the DOH's website. Other sources of anonymized aggregated data can come from Electronic Medical Record (EMR) Systems being used by different private and public hospitals, if given access. The data collected, along with the spatio-temporal details (geo-location and timestamp) embedded in it, can serve as a baseline for possible correlation with the collected Twitter posts to make an infodemiological model. Using these data, the goal of this research is to produce a model that shows and predicts the conditions of various locations regarding the chosen diseases.

% As the main government entity in charge of public health, the DOH, through the epidemiology bureau, 
%This research focuses on the clusters of people in the Philippines, having a municipality as the smallest possible collective unit. 


The research is divided into two phases, which are 1) creation of a twitter collection system for health related topics and visualization on a map, and 2) the modeling and prediction of possible areas of outbreaks based on Twitter posts. The major part of the initial phase is in the collection, pre-processing, quantification, and visualization of related tweets. This phase's primary concern is creating a listener for health related words relevant to infodemiology, such as symptoms and associated outbreaks. Users' sentiments will  disregarded, as the model to be made focuses on classifying whether a tweet is infodemiological in nature or not. Bilingual keywords or phrases will be collected, which is vital since the Philippines uses two primary languages, the Filipino language and the English language. The second phase is where the modeling of the collected tweets will happen. An infodemiological model will be developed to identify possible locations of outbreaks based on current tweets through IBM's Spatio-Temporal Epidemiological Modeler.

Social networking websites had already been used by researchers to track different diseases like Human immunodeficiency virus (HIV) and Dengue \cite{Young:2014aa, gomide2011dengue}. This research would follow some methodologies used by previous studies, but this would cater more to the Philippines since the data comes from the Philippines' domain. Added to this, the factor of bilinguality may add more needed appropriations to the study since Filipinos are highly likely to use social networking websites in both English and in Filipino.

\input{ResearchQuestions}
\input{SignificanceOfTheStudy}